Manuscript draft

### 2024-04-05 docs/manuscript.qmd
updated with a brief description of the degree and strength z-scores

## Summarizing movement measures
```{r}
#| include: false
load(here("data/calcMovement/pca_space.Rda"))
propvar_space <- as.data.frame(summary(pca_space)$importance)$PC1[2]

@@

### Co-roosting 
Vultures were considered to be co-roosting if they roosted within 500m of each other on a given night. Unlike interactions in the co-feeding and co-flight networks, co-roosting interactions could occur at a maximum frequency of once per night. Therefore, edge weights in the co-roosting network represented the number of nights on which two individuals co-roosted divided by the number of nights when both individuals had a valid roost location (i.e. they theoretically could have co-roosted), regardless of proximity.

@@

We used the R package {igraph} [@csardi2005] to calculate each individual's degree and strength for each social situation in each season. Degree is defined as the number of other individuals that a vulture is connected to (i.e. ever interacted with). Strength measures how much an individual interacts, and is calculated as the sum of the weights of all edges that the focal vulture is involved in.

@@

```{r}
#| label: fig-d-space-situ
#| fig-cap: "Relationship between normalized degree and space use, with separate trend lines for each social situation"
plot_d_eff_situ_space
plot_d_emt_situ_space
```
There was a significant relationship between how vultures used space and how many other individuals they encountered in flight. Vultures that consistently ranged widely encountered more others in flight, while those that concentrated their movement in smaller home ranges and smaller proportional core areas flew with fewer others. There was a slight, albeit non-statistically-significant, trend toward lower degree in the co-feeding and co-roosting networks for vultures with larger ranges.

@@

```{r}
#| eval: false
#| label: fig-s-space
#| fig-cap: "Relationship between normalized strength and space use"
plot_s_eff_situ_space
plot_s_emt_situ_space
```
Vultures with greater space use had significantly more co-flight interactions (higher strength in the co-flight network). They also tended to have a higher co-roosting strength, though the trend was not statistically significant. There was no relationship between seasonal space use and co-feeding strength.

>>>>

## Summarizing movement measures

```{r}
#| include: false
load(here("data/calcMovement/pca_space.Rda"))
propvar_space <- as.data.frame(summary(pca_space)$importance)$PC1[2]

@@

### Co-roosting

Vultures were considered to be co-roosting if they roosted within 500m of each other on a given night. Unlike interactions in the co-feeding and co-flight networks, co-roosting interactions could occur at a maximum frequency of once per night. Therefore, edge weights in the co-roosting network represented the number of nights on which two individuals co-roosted divided by the number of nights when both individuals had a valid roost location (i.e. they theoretically could have co-roosted), regardless of proximity.

@@

We used the R package {igraph} [@csardi2005] to calculate each individual's degree and strength for each social situation in each season. Degree is defined as the number of other individuals that a vulture is connected to (i.e. ever interacted with). Strength measures how much an individual interacts, and is calculated as the sum of the weights of all edges that the focal vulture is involved in. We normalized degree by dividing it by the number of individuals present in the network in the current season minus one. We normalized strength by dividing it by the sum of all strength values in the network in the current season \[REF\].

## Comparison to a reference model

Since social interactions can arise from both spatial proximity and social preference (i.e. "passive" versus "active" interactions [@smith-aguilar2016]). We can use spatially explicit data-stream randomization to compare individuals' observed interaction rates to those expected to arise from chance encounters by individuals sharing space without synchronizing their movements [@gahm; @spiegel2016]. This allows us to decouple social and spatial drivers of social interactions. It also helps us account for how individuals' locations in space (e.g. the locations of their home ranges) affect how many social interactions they end up having.

Using the "wrap-around" method outlined in [@gahm], we conducted 100 iterations of the wrap-around randomization for each season of data separately, using a maximum time-shift of 1 day either backward or forward. We calculated normalized degree and strength for each individual in each season in each of the 100 replicate randomizations. Then, we compared each individual's observed normalized degree and strength to the distribution of 100 normalized degree and strength values obtained from the randomization replicates and calculated a z-score for the observed value to measure the extent to which it deviated from random. These z-scores were used as the response variables in our mixed models.

@@

```{r}
#| label: fig-d-space-situ
#| fig-cap: "Relationship between normalized degree and space use, with separate trend lines for each social situation"
plot_d_eff_situ_space
plot_d_emt_situ_space
```

There was a significant relationship between how vultures used space and how many other individuals they encountered in flight. Vultures that consistently ranged widely encountered more others in flight, while those that concentrated their movement in smaller home ranges and smaller proportional core areas flew with fewer others. There was a slight, albeit non-statistically-significant, trend toward lower degree in the co-feeding and co-roosting networks for vultures with larger ranges.

@@

```{r}
#| eval: false
#| label: fig-s-space
#| fig-cap: "Relationship between normalized strength and space use"
plot_s_eff_situ_space
plot_s_emt_situ_space
```

Vultures with greater space use had significantly more co-flight interactions (higher strength in the co-flight network). They also tended to have a higher co-roosting strength, though the trend was not statistically significant. There was no relationship between seasonal space use and co-feeding strength.
